---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr)
data <- read.csv("activity.csv")
str(data)
```

## What is mean total number of steps taken per day?

```{r}
count_step <- data %>% group_by(date) %>% summarise(count = sum(steps, na.rm = TRUE))
hist(count_step$count, main = "total number of steps/day", xlab = "number of steps", las = 1, col = "grey80")
```


The **mean** and **median** total number of steps taken per day are `r mean(count_step$count, na.rm = T)` and `r median(count_step$count, na.rm = T)`, respectively.


## What is the average daily activity pattern?
```{r}
step_inter <- data %>% group_by(interval) %>% summarise(mean = mean(steps, na.rm = TRUE))
plot(step_inter, type = "l", las = 1, col = "darkblue")
```

The maximum number of steps occurs at `r step_inter$interval[which(max(step_inter$mean) == step_inter$mean)]` 5-minute interval with `r max(step_inter$mean)` steps.

## Imputing missing values
The total number of missing values in the dataset is `r sum(is.na(data$steps))`. This corresponds to `r sum(is.na(data$steps))/nrow(data)*100`% the total entrances.

Create a new dataset using mean for 5-minute interval:

```{r}
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
mean_day <- data %>% group_by(interval) %>% mutate(steps = impute.mean(steps))
```

```{r}
count_step2 <- mean_day %>% group_by(date) %>% summarise(count = sum(steps, na.rm = TRUE))
hist(count_step2$count, main = "total number of steps/day", xlab = "number of steps", las = 1, col = "grey80")
```

The **mean** and **median** total number of steps taken per day are `r mean(count_step2$count, na.rm = T)` and `r median(count_step2$count, na.rm = T)`, respectively.

These values differ (`r mean(count_step2$count, na.rm = T)` and `r median(count_step2$count, na.rm = T)`) from the estimates from the first part of the assignment (`r mean(count_step$count, na.rm = T)` and `r median(count_step$count, na.rm = T)`). When the missing values were replaced by mean values, the shape of histogram became more normal distributed.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
mean_day$date2 <- weekdays(as.Date(mean_day$date))
weekday <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
mean_day$date3 <- factor(mean_day$date2 %in% weekday, levels=c(F, T), labels = c('weekend', 'weekday'))

step_inter2 <- mean_day %>% group_by(date3, interval) %>% summarise(mean = mean(steps, na.rm = TRUE))
```

```{r}
library(lattice)
xyplot(mean ~ interval|date3, data = step_inter2, type = "l", layout = c(1,2), ylab = "Average of number of steps")
```


